### 2019-01-10 paper/SI.tex
Change SI for delta value part

\documentclass[manuscript=suppinfo,email=true, hyperref=true, keywords=false]{achemso}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{times}
\usepackage{todonotes}
\newcommand{\Fig}{Fig.}
\renewcommand{\thesection}{S\arabic{section}}
\SectionNumbersOn
\usepackage{xr-hyper}
\externaldocument[main-]{paper}

% \usepackage{lineno}
% \linenumbers

\author{Roman M. Wyss} \affiliation{Soft Materials,
  Department of Materials, Eidgenössische Technische
  Hochschule (ETH) Zürich, Vladimir-Prelog-Weg 1-5, Zürich CH-8093,
  Switzerland.  }
\altaffiliation{R. M. S. and T. T. contributed equally to this work}
\author{Tian Tian} \affiliation{Institute for
  Chemical and Bioengineering Department of Chemistry and Applied
  Biosciences, Eidgenössische Technische Hochschule (ETH) Zürich,
  Vladimir-Prelog-Weg 1-5, Zürich CH-8093, Switzerland.}
\altaffiliation{R. M. S. and T. T. contributed equally to this work}

\author{Khadija Yazda} \affiliation{Nanoscience for Energy Technology and Sustainability,
Department of Mechanical and Process Engineering,
Eidgenössische Technische Hochschule (ETH) Zürich,
Tannenstrasse 3, Zürich CH-8092, Switzerland.}

\author{Hyung Gyu Park} \affiliation{Nanoscience for Energy Technology and Sustainability,
Department of Mechanical and Process Engineering,
Eidgenössische Technische Hochschule (ETH) Zürich,
Tannenstrasse 3, Zürich CH-8092, Switzerland.}
\alsoaffiliation{Mechanical Engineering, Pohang University of Science and
  Technology (POSTECH), 77 Cheongam-ro, Nam-gu, Pohang, Gyeongbuk,
  Republic of Korea.}
\email{iduserpark@gmail.com}

\author{Chih-Jen Shih}  \affiliation{Institute for
  Chemical and Bioengineering Department of Chemistry and Applied
  Biosciences, Eidgenössische Technische Hochschule (ETH) Zürich,
  Vladimir-Prelog-Weg 1-5, Zürich CH-8093, Switzerland.}
\email{chih-jen.shih@chem.ethz.ch}

\title{Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene}

%%%%%%%%%%%%%%%%%%%%
%Main
%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}

\section{Graphene membrane manufacturing and results.}
\label{sec:exp}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/SI-1.png}
  \caption{\textbf{a}. Schematic of manufacturing process for
    patterning of double layer graphene. Graphene is synthesized and
    transferred to a substrate (glass slide) to yield a double
    layer. s-BCP is spin-coated and annealed in vacuum to undergo microphase
    separation, followed by plasma etch and wet etching yielding a
    porous polystyrene (PS) mask. Anisotropic etching of the mask
    leads to patterning the underlying graphene, PS is removed by
    thermal annealing. The porous graphene is then transferred to a
    substrate (e.g. PCTE). \textbf{b}. Raman spectra of graphene
    before (orange curve) and after (blue
    curve) patterning. An increase of the D-peak at 1350 cm$^{-1}$ indicates the
    formation of defects and edges. The spectra have been normalized
    to the G-peak intensity and offset for better
    comparison. \textbf{c}. SEM graph of patterned graphene on PCTE at
    lower (top) and higher (bottom) magnifications.
    % Note that tears in
    % graphene appear bright white with a black center.
  }
  \label{fig:exp}
\end{figure}

\section{Ion diffusion measurement and calculation}
\label{sec:measure}

\begin{table}[htbp]
  \centering
  \begin{tabular}{lc}
    \hline
    Salt & $D$ (10$^{9}$ m$^{2}\cdot$s$^{-1}$) \\
    \hline
    K$^{+}$     & 1.95 \\
    Na$^{+}$    & 1.33 \\
    Li$^{+}$    & 1.03 \\
    SO$_{4}^{2-}$       & 1.06 \\
    Ca$^{2+}$   & 0.79\\
    Mg$^{2+}$ & 0.7\\
    Fe{[(CN)$_{6}$]}$^{3-}$   & 0.90\\
    Cl$^{-}$ &  2.03 \\
               \hline
  \end{tabular}
  \caption{Different diffusivities for the anions and cations of the
    salts used in this work \cite{vanysek_2000}. The theoretical
    diffusion through PCTE can be calculated using Eq. 1 in the main
    text and the diffusivities of the respective anion and cation.}
  \label{tab:diff}
\end{table}

The ion diffusion through b-PCTE has been performed using a sample
sandwiched between two layers of Aluminum tape where the edges have
been sealed using water-resistant epoxy (ACS Marine Epoxy) to avoid
interlayer leakage pathways \cite{Choi_2018}.  The sample was inserted in
the fixture following the same wetting procedure described in the main
text Methods and was kept wet for the entire series of experiments.

The conductivity increases $s$ (mS$\cdot$cm$^{-1}\cdot$d$^{-1}$) is
converted into ion permeation rates $\boldsymbol{J}_{\mathrm{PCTE}}$
(mol$\cdot$m$^{-2}\cdot$s$^{-1}$) by use of a calibration factor
$c_{\mathrm{f}}$, the volume of LCR ($V_{\mathrm{LCR}}$) , the area of
the membrane $A$ and the time $t$ as:
\begin{equation}
  \label{eq:JPCTE}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{s V_{\mathrm{LCR}}}{c_{\mathrm{f}} A t}
\end{equation}
The calibration factors for all salts have been obtained by a linear
fit of conductivity versus concentration for 4 solutions with
concentrations from 10$^{-4}$ M to 10$^{-2}$ M and extracting the
resulting slope. The rest of the factors are $V_{\mathrm{LCR}}$ = 7.33
mL, $A$ = 9$\times$10$^{-6}$ m$^{-2}$. Table \ref{tab:exp-2} shows the
averaged conductivity increase s, the extracted calibration factors
$c_{\mathrm{f}}$ and the resulting ion permeation rates using
Eq. \ref{eq:JPCTE}. The values from this table are shown in main text \Fig 2b
(blue bars).
\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    Salt & $s$ (mS$\cdot$cm$^{-1}$d$^{-1}$) & $c_{\mathrm{f}}$ (mS$\cdot$cm$^{-1}\cdot$mol$^{-1}\cdot$ml) &
                                                                                                            $\boldsymbol{J}_{\mathrm{PCTE}}$ (10$^{6}$ mol$\cdot$m$^{-2}\cdot$s$^{-1}$)\\
    \hline
    KCl & 1.51$\pm$0.27$\times$10$^{-2}$  & 106   & 1.34 $\pm$ 0.24\\
    NaCl        & 9.90$\pm$0.17$\times$10$^{-3}$ & 91    & 1.02 $\pm$ 0.02\\
    LiCl        & 6.03 $\pm$0.80$\times$10$^{-3}$        & 83    & 0.682 $\pm$ 0.090\\
    CaCl$_{2}$  & 1.25$\pm$0.1$\times$10$^{-2}$    & 149   & 0.789 $\pm$ 0.063\\
    MgSO$_{4}$  & 8.80$\pm$1.2$\times$10$^{-3}$   & 130   & 0.633 $\pm$ 0.084\\
    K$_{3}$[Fe(CN)$_{6}]$ & 2.40$\pm$0.28$\times$10$^{-2}$   & 395   & 0.575 $\pm$ 0.063\\
    K$_{2}$SO4$_{4}$    & 1.98$\pm$0.20$\times$10$^{-2}$  & 258   & 0.722 $\pm$ 0.073\\
    \hline
  \end{tabular}
  \caption{Extracted slope $s$ as an averaged result from 3 consecutive
    measurements, the calibration factors $c_{\mathrm{f}}$ and the resulting ion
    permeation $\boldsymbol{J}_{\mathrm{PCTE}}$ through b-PCTE membranes for all salts.}
  \label{tab:exp-2}
\end{table}

\section{Control experiment with copper tape}
\label{sec:copper}
We make sure that the applied voltage is effectively applied via the
graphene membrane and not through leakage directly coupling the
coppertape to the ionic solution. For this, we use a device as in \Fig
1a, omitting the graphene and measuring the current through the
membrane with the Autolab electrochemical workstation using the
membrane as working electrode and a platinum wire as counter/reference
electrode 3.5 cm apart in 0.1 mM KCl. When no graphene is inserted in
the device, a current in the baseline range ($\sim$ 0.1 nA) is
observed, indicating no current passing from the membrane. However, if
graphene is inserted, a current in the range of $\sim$ 10 - 100 nA is
measured. If the copper tape is in contact with the solution when
omitting the Kapton tape, the current increases to $\sim$ 1 - 10
$\mathrm{\mu}$A, independent of the exposed area of the copper tape to
the solution.

\section{Circuit analog model of transport through PG-PCTE}
\label{sec:R-model}
The ionic transport through the PG-PCTE membrane is the concerted
effect the transport resistance through PG and PCTE membranes,
respectively. As schematically shown in \Fig{} \ref{fig:R-model}, the
overall transport resistance of the system is composed of the series
resistance $R_{\mathrm{G}}$ and $R_{\mathrm{PCTE}}$, respectively. We
further treat $R_{\mathrm{G}}$ as the effective resistance of $\gamma$
parallel resistors with individual resistance $R_{\mathrm{g}}$, where
$\gamma$ is the average numbers of PG pores per PCTE channel, and
$R_{\mathrm{g}}$ is the average transport resistance per PG pore. 
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-resistance-rectification.pdf}
  \caption{Simple circuit analog model for the transport through
    PG-PCTE membrane. \textbf{a}. Schematic of PG-PCTE membrane in
    cross-section view. \textbf{b}. Model for transport resistance in
    analogy to an electric circuit. This schematic depicts a cut
    through a PG-PCTE membrane showing a single PCTE pore and multiple
    graphene pores.}
  \label{fig:R-model}
\end{figure}
We then estimate the value of $\gamma$ from geometric analysis of both
membranes. Given the pore number density of PG
($\sim$1.25$\times$10$^{14}$ m$^{-2}$), the pore number density of
PCTE ($\sim$1.50$\times$10$^{12}$ m$^{-2}$), and the percentage of
PCTE pores with PG on top ($\sim$0.13), $\gamma$ is determined to be
$\sim$10.83.

Next we seek the value between $R_{\mathrm{PCTE}}$ and
$R_{\mathrm{G}}^{0}$ (ungated resistance of PG), $\delta$ (as defined
in main text Section \ref{main-sec:res-1}). It is pointed out that the
scaling law for $R_{\mathrm{PCTE}}$ and $R_{\mathrm{G}}^{0}$
differs. For a long channel like PCTE, the resistance is derived from
Fick's law as\cite{O_Hern_2014}:
\begin{equation}
\label{eq:R-PCTE}
 R_{\mathrm{PCTE}} = \frac{1}{D} \frac{L}{\pi r_{\mathrm{PCTE}}^{2}}
\end{equation}
whereas $D$ is the diffusivity of the specie, $L$ is the length of
channel ($\sim$24 $\mathrm{\mu}$m), and $r_{\mathrm{PCTE}}$ is the
average radius of the PCTE channels ($\sim$400 nm). On the other hand,
the transport through 2D pore with high radius/thickness ratio falls
into the regime of the Sampson formula \cite{Roscoe_1949}, where the
entrance resistance dominates over the bulk resistance, with the
formula as \cite{Roscoe_1949,Celebi_2014}:
\begin{equation}
  \label{eq:sampson}
  R_{\mathrm{g}}^{0} = \frac{1}{D} \frac{3 \pi}{r_{\mathrm{G}}}
\end{equation}
We therefore estimate the value of $\delta$ as:
\begin{equation}
  \label{eq:delta-resistance}
  \delta = {\displaystyle
    \frac{{\displaystyle \frac{L}
        {\pi r_{\mathrm{PCTE}}^{2}}}}
    {{\displaystyle \frac{3 \pi}{r_{\mathrm{G}}} \frac{1}{\gamma}}}}
  \approx 2.20
\end{equation}

\section{KCl Debye length and conductivity}
\label{sec:debye}
Table \ref{tab:debye} gives an overview about the measured
conductivities and their calculated Debye length values for KCl. The
information is used to plot the $\xi$ versus $c_{0}$ curve.

\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    $c_{0}$ (mM) & Conductivity (mS$\cdot$cm$^{-1}$) & $\lambda_{\mathrm{D}}$ (nm)\\
    \hline
    0.1&        1.90$\times$10$^{-2}$ &        30.4\\
    0.33&       5.41$\times$10$^{-2}$ & 16.0\\
    1   &1.33$\times$10$^{-1}$ & 10.0\\
    3.3&     5.13$\times$10$^{-1}$ &  6.0\\
    10&      1.08    &3.3\\
    \hline
  \end{tabular}
  \caption{Conductivity and calculated Debye lengths of various KCl
    concentrations in order to obtain rectification versus
    concentration curves. }
  \label{tab:debye}
\end{table}

\section{Measurement of charge carrier density of patterned graphene}
\label{sec:charge-dens}
The intrinsic charge carrier density of graphene was measured using a
field electron transistor (FET) in air with a 300nm SiO$_2$ as a
gate. The device setup is shown in \Fig{} \ref{fig:charge-dens}a,
where the source and drain electrodes were deposited onto the PG
sample transfered onto 300nm SiO$_2$ on doped Silicon. The
characteristic drain-source current $I_{\mathrm{DS}}$ as a function of
$V_{\mathrm{G}}$ is shown in \Fig{} \ref{fig:charge-dens}. The
charge-neutral point gate voltage $V_{\mathrm{CNP}}$ is measured to be 2.8 V.
Using the capacitance of the SiO$_{2}$ dielectric, we an intrinsic
charge density of $\sim{}$1.4$\times$10$^{11}$
\textit{e}$\cdot$cm$^{-2}$, and a shift in the PG's Fermi level by
$\sim{}$48 meV, which is negligible compared with the gate-induced
charge density in our experiments. Therefore, the assumption that the
PG sample is intrinsic without gating, is justified.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-gr-transistor.pdf}
  \caption{Characterization of the intrinsic charge density of the PG
    sample using field effect transistor. \textbf{a}. The setup of the
    PG-FET. \textbf{b}. The I$_{\mathrm{DS}}$ as a function of
    $V_{\mathrm{G}}$. The charge neutral point gate voltage is
    determined as $\sim{}$2.8 V, corresponding to an intrinsic charge
    density of $\sim{}$1.4$\times$10$^{11}$
    \textit{e}$\cdot$cm$^{-2}$, which is negligible compared with the
    gate-induced charges.}
  \label{fig:charge-dens}
\end{figure}

\section{Further details of the numerical simulation }
\label{sec:numer}

The numerical simulations were carried out using COMSOL Multiphysics
5.3a. The simulation domain is shown in \Fig \ref{fig:scheme}. To
simplify the geometry, we use axial symmetric coordinate system. The
radius \textit{L} and height \textit{H} of both HCR and LCR are set to
20 $R_{\mathrm{G}}$, the radius of the nanopore. The potential at the
bottom of the LCR and the center of the graphene domains are set to be
0 and $\psi_{\mathrm{G}}$, respectively. The potential at the end of
HCR is not set explicitly, but rather determined through the Poisson
equation with no flux boundary condition
$\boldsymbol{D}^{\mathrm{norm}}=0$. The relative permittivity in the
HCR and LCR is set as $\varepsilon_{\mathrm{}}=78$. The relative
permittivity of the Stern layer is set to
$\varepsilon_{\mathrm{H}}=20$\cite{Conway_1951} and for the graphene
domain we set the relative permittivity as a large number
(e.g. $10^{5}$) for a better convergence of the solver.  For the
transport of diluted ionic species, we use the PNP equations to
describe the diffusion and drift of the ions, which can be further
unified by the gradient of electrochemical potential, as described in
the main text. The initial values of the electrolyte is set to $c_{0}$
in HCR and 0.1 $c_{0}$ in LCR. At the ends of the cells in the
z-direction, we use the inlet boundaries conditions for the
concentration of electrolytes, that the concentration on these
boundaries remains the same with the initial value, due to the
effective mixing in the experimental setup.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-numerical-1.png}
  \caption{Scheme of the simulation domain used. Left: geometry of the
    whole simulation domain. Right: configuration near the graphene
    nanopore, corresponding to the red rectangle on the left side.}
  \label{fig:scheme}
\end{figure}

The transport of ionic species is described by the steady-state
Nernst-Planck equation:
\begin{eqnarray}
  \label{eq:pnp}
  \nabla \cdot \boldsymbol{J}_{i} = -\nabla \cdot \left( \frac{D_{i}}{k_{\mathrm{B}}T} c_{i} N_{\mathrm{A}} \nabla \mu_{i}\right)
  = -\nabla \cdot \left( \frac{D_{i}}{k_{\mathrm{B}} T} c_{i} N_{\mathrm{A}}
  [k_{\mathrm{B}}T \nabla \ln x_{i} + z_{i} e \nabla \psi]\right) = 0\\
  x_{i} = \frac{c_{i}}{\sum_{i} c_{i} + c_{\mathrm{H_{2}O}}}
\end{eqnarray}
The potential distribution within the whole simulation domain is
described by the Poisson equation:
\begin{equation}
  \label{eq:poisson}
  \nabla \cdot (\varepsilon_{\mathrm{m}} \varepsilon_{0} \nabla \psi) = -N_{\mathrm{A}} e \sum_{i} c_{i}z_{i}
\end{equation}
Where $\varepsilon_{\mathrm{m}}$ is the relative permittivity of
domain m. The contribution of mobile charges is only valid in the
solution domain. The use of a large $\varepsilon_{\mathrm{G}}$ ensures
that the potential in the whole simulation domain can be solved
continuously.

To simulate the applied external bias $V_{\mathrm{G}}$, we set the
value of $\psi_{\mathrm{G}}$ explicitly and extract the charge
$\sigma_{\mathrm{G}}$ by:
\begin{equation}
  \label{eq:sigma-G}
  \sigma_{\mathrm{G}} = - {\displaystyle \frac{\int_{\mathrm{\Omega}} z_{i} c_{i} N_{\mathrm{A}} e \mathrm{d}^{3} \Omega}{S_{\mathrm{G}}}}
\end{equation}
and $V_{\mathrm{G}}$ is then calculated by:
\begin{eqnarray}
  \label{eq:VG}
  V_{\mathrm{G}} = \Delta \phi_{\mathrm{G}} + \psi_{\mathrm{G}}\\
  \sigma_{\mathrm{G}} = \int_{0}^{\Delta \phi_{\mathrm{G}}} \frac{1}{C_{\mathrm{Q}}(\phi_{\mathrm{G}})} \mathrm{d} \phi_{\mathrm{G}}
\end{eqnarray}
where the quantum capacitance
$C_{\mathrm{Q}}=\partial \sigma_{\mathrm{G}}/\partial
\phi_{\mathrm{G}}$ is calculated by Eq. 8 in main text.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\linewidth]{img/SI-trap.png}
  \caption{Schematic illustration of the origin of the asymmetric
    rectification with respect to $V_{\mathrm{G}}$. Due to the
    existence of surface charge traps at negative $V_{\mathrm{G}}$,
    the graphene is less charged compared to the positive
    $V_{\mathrm{G}}$.}
  \label{fig:trap}
\end{figure}

Note that due to the existence of electron traps on graphene due to
fabrication process, the induced charge traps on graphene effectively
reduce the charge density on graphene and greatly attenuate the
surface potential on graphene, as illustrated in \Fig
\ref{fig:trap}. This effect is universally observed in
CVD-graphene-based field effect transistors in air, while the
mechanism is fully understood yet. In view of this, we only simulate
the situation where $V_{\mathrm{G}}>$ 0 for the salt rejection
mechanism. The trend that $\xi$ is almost independent of
$V_{\mathrm{G}}$ when $V_{\mathrm{G}}<$ 0 can be qualitatively
explained by the existence of surface charge traps as stated above.
We have also tested the mesh-independency of our solutions. As show in
in \Fig{} \ref{fig:mesh}a, we use two parameters to control the
refinement of the mesh entities: (i) trangle mesh division
$N_{\mathrm{div}}$, giving the smallest triangle mesh size
$\delta_{\mathrm{G}} / N_{\mathrm{div}}$, where $\delta_{\mathrm{G}}$
is the thickness of graphene, and (ii) division of mesh
$N_{\mathrm{b}}$, on the boundary $0<r<R_{\mathrm{G}};z=0$. The
solution convergence is evaluated by the mean absolute error (MAE) of
anion flux $\boldsymbol{J}_{z}$ on the boundary
$0<r<R_{\mathrm{G}};z=0$, as shown in \Fig{}
\ref{fig:mesh}b. The solution of the PNP model is sensitive to the
value $N_{\mathrm{b}}$, and can be regarded numerically converged when
$N_{\mathrm{b}}>100$. Combining the calculation effort and numerical
precision, we adapt the value $N_{\mathrm{div}}=5$ and
$N_{\mathrm{b}}=100$ in our simulations.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-mesh.pdf}
  \caption{Mesh independence of the solution. \textbf{a}. Scheme of
    the mesh refinement near the graphene pore. The values
    $N_{\mathrm{div}}$ and $N_{\mathrm{b}}$ controls the size of mesh
    along the boundary $z=0, 0<r<R_{\mathrm{G}}$ (shown in
    red). \textbf{b}. Mean absolute error (MAE) of the anion flux
    $\boldsymbol{J}_{z}$ along the boundary $z=0,
    0<r<R_{\mathrm{G}}$. The error is more sensitive to
    $N_{\mathrm{b}}$ than $N_{\mathrm{div}}$.}
  \label{fig:mesh}
\end{figure}

\section{Further numerical simulation results}
\label{sec:simu-res}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-concentration.png}
  \caption{Cation concentration $c_{+}$ (\textbf{a}.) and anion
    concentration $c_{-}$ (\textbf{b}.) near the graphene nanopore for
    KCl solution at 0.1 mM. Depletion of cation and accumulation of
    anion can be observed near the graphene surface.}
  \label{fig:conc}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-electrochemical-decomposite.pdf}
  \caption{Decomposition of electrochemical potential near a nanopore
    corresponding to \Fig 4 in main text. \textbf{a}. The electrostatic
    (drift) contribution to the electrochemical potential for cation
    (left) and anion (right) and \textbf{b}. The concentration (diffusion)
    contribution to the electrochemical potential for cation (left)
    and anion (right). The relative small order of magnitude for anion
    electrochemical potential is caused by the balance between the
    diffusion and drift for anion.}
  \label{fig:potential}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-flux.png}
  \caption{z-component flux $\boldsymbol{J}_{z}$ of cation
    (\textbf{a}.)  and anion (\textbf{b}.) near the graphene nanopore
    corresponding to Figure 4b in main text. The pathways for the
    different ions can be spatially distinguished: the flux of cation
    passes mostly through the center of the graphene nanopore while
    the anion passes mainly through the edge. The analysis is in
    consistent with the gradient of chemical potential as shown in the
    main text \Fig 4b.}
  \label{fig:flux}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\linewidth]{img/SI-rect-reverse.pdf}
  \caption{Average total flux $\boldsymbol{J}_{z}$ in the nanopore as
    a function of $\psi_{\mathrm{G}}$ at different concentrations. The
    range of $\psi_{\mathrm{G}}$ is larger than the experimentally
    achievable value. The rectification of flux becomes weaker after a
    certain level of $\psi_{\mathrm{G}}$ is reached. Further
    increasing the surface potential of graphene may even enhance the
    transportation of ions, which is similar to the cases of ionic
    transistors.}
  \label{fig:reverse}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-flux-larger.png}
  \caption{Spatial distribution of $\boldsymbol{J}_{z}$ as a function
    of $r$ in a 20 nm nanopore at different levels of
    $\psi_{\mathrm{G}}$ (larger than experimentally achievable values)
    for cation (\textbf{a}.) and anion (\textbf{b}.). The cation flux
    becomes nearly saturated with increasing $\psi_{\mathrm{G}}$,
    while the flow of anion near the pore edge becomes much
    larger. The drastic increase of pore-bounded anion flux is caused
    by the accumulation of anions near the graphene surface, and in
    turn causes the increasing of total flux as seen in \Fig
    \ref{fig:reverse}.}
  \label{fig:large-V}
\end{figure}
\clearpage{}

\section{Conductivities and Debye length for all salts}
\label{sec:salts}
In Table \ref{tab:debye}, the conductivities and calculated Debye
lengths for 6 salts are shown. At the same concentration, the higher
ionic strength of the multivalent salts leads to a reduction of the
respective Debye lengths.

\begin{table}[htbp]
  \centering
  \begin{tabular}{lcc}
    \hline
    Solution & Conductivity (mS$\cdot$cm$^{-1}$) & $\lambda_{\mathrm{D}}$ (nm) \\
    \hline
    0.1 mM NaCl &1.80$\times$10$^{-2}$  &30.4\\
    0.1 mM LiCl &1.58$\times$10$^{-2}$ &30.4\\
    0.1 mM CaCl$_{2}$&  2.19$\times$10$^{-2}$ &17.2\\
    0.1 mM MgSO$_{4}$   &3.27$\times$10$^{-2}$ &15.2\\
    0.1 mM K$_{2}$SO$_{4}$      &4.16$\times$10$^{-2}$ &17.2\\
    0.1 mM K$_{3}$[Fe(CN)$_{6}$]&       5.95$\times$10$^{-2}$  &12.4\\
    \hline
  \end{tabular}
  \caption{Conductivity and $\lambda_{\mathrm{D}}$ of various salts.}
  \label{tab:conductivity}
\end{table}

\section*{}
\label{sec:ref}
\bibliography{ref}

\end{document}