### 2019-01-15 paper/paper.tex
Remove blank line

\documentclass[journal=langd5,email=true, hyperref=true, keywords=false]{achemso}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{times}
\usepackage{todonotes}
\DeclareCaptionLabelSeparator{bar}{  \textbar\quad}
\usepackage[labelsep=bar, labelfont=bf]{caption}
\newcommand{\Fig}{Fig.}
\usepackage{lineno}
\usepackage{xr-hyper}
\externaldocument[SI-]{SI}

\author{Roman M. Wyss} \affiliation{Soft Materials,
  Department of Materials, Eidgenössische Technische
  Hochschule (ETH) Zürich, Vladimir-Prelog-Weg 1-5, Zürich CH-8093,
  Switzerland.  }
\altaffiliation{R. M. W. and T. T. contributed equally to this work}
\author{Tian Tian} \affiliation{Institute for
  Chemical and Bioengineering Department of Chemistry and Applied
  Biosciences, Eidgenössische Technische Hochschule (ETH) Zürich,
  Vladimir-Prelog-Weg 1-5, Zürich CH-8093, Switzerland.}
\altaffiliation{R. M. W. and T. T. contributed equally to this work}

\author{Khadija Yazda} \affiliation{Nanoscience for Energy Technology and Sustainability,
Department of Mechanical and Process Engineering,
Eidgenössische Technische Hochschule (ETH) Zürich,
Tannenstrasse 3, Zürich CH-8092, Switzerland.}

\author{Hyung Gyu Park} \affiliation{Nanoscience for Energy Technology
  and Sustainability, Department of Mechanical and Process
  Engineering, Eidgenössische Technische Hochschule (ETH) Zürich,
  Tannenstrasse 3, Zürich CH-8092, Switzerland.}
\alsoaffiliation{Mechanical Engineering, Pohang University of Science and
  Technology (POSTECH), 77 Cheongam-ro, Nam-gu, Pohang, Gyeongbuk,
  Republic of Korea.}
\email{iduserpark@gmail.com}

\author{Chih-Jen Shih}  \affiliation{Institute for
  Chemical and Bioengineering Department of Chemistry and Applied
  Biosciences, Eidgenössische Technische Hochschule (ETH) Zürich,
  Vladimir-Prelog-Weg 1-5, Zürich CH-8093, Switzerland.}
\email{chih-jen.shih@chem.ethz.ch}

\title{Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene}

%%%%%%%%%%%%%%%%%%%%
%Main
%%%%%%%%%%%%%%%%%%%%
\begin{document}
\clearpage{}
\begin{abstract}
  {\normalsize Atomically thin porous graphene membranes areis
    emerging as one of the most promising candidates for the
    next-generation membrane technologymaterial owing to the ultrahigh
    permeation. However, the transport selectivity relies on the
    precise control over pore size and shape, which considerably
    limitingcompromises the scalability. Here, we study electrolyte
    permeation through a sheet of large-area, porous graphene
    membrane, with relatively large pore sizes of 20$\pm$10
    nanometers. Counterintuitively, we observe macroscopica high
    degree of salt rejection is observed by applying
    electrostaticallyelectrostatic gating, which reducesreducing the
    diffusive flux by up to one order of magnitude. We systematically
    investigate the effects of salt concentration and species,
    including developing a theory to model the electrolyte diffusion
    through a nanopore drilled in a sheet of gated graphene. The
    interplay between graphene quantum capacitance and the electrical
    double layer is found to selectively modulate the anionic and
    cationic transport paths, creating voltage-dependent
    electrochemical barriers when the pore size is comparable to the
    Debye length. Our findings reveal a new degree of freedom
    regulating electrolyte permeation through porous two-dimensional
    materials.
    % Atomically thin porous graphene is emerging as one of
    % the most promising candidates for the next-generation membrane
    % technology owing to the ultrahigh permeation. However, the
    % transport selectivity relies on the precise control over pore size
    % and shape, considerably limiting the scalability. Here, we study
    % electrolyte permeation through a sheet of large-area, porous
    % graphene membrane, with relatively large pore sizes of 20$\pm$10
    % nanometers. Counterintuitively, we observe macroscopic salt
    % rejection by applying electrostatically gating, which reduces the
    % diffusive flux by up to one order of magnitude. We systematically
    % investigate the effects of salt concentration and species,
    % including developing a theory to model the electrolyte diffusion
    % through a nanopore in a sheet of gated graphene. The interplay
    % between graphene quantum capacitance and the electrical double
    % layer is found to selectively modulate the anionic and cationic
    % transport paths, creating voltage-dependent electrochemical
    % barriers when the pore size is comparable to the Debye length. Our
    % findings reveal a new degree of freedom regulating electrolyte
    % permeation through porous two-dimensional materials.   
    % Atomically thin porous graphene membranes are emerging
    % as one of the most promising candidates as next-generation
    % membrane technology owing to their unique transport physics and
    % high permeation. Among the current challenges of 2D membranes is
    % the understanding of fundamental ion transport as well as its
    % control towards applications in ion rectification, ionic
    % transistors or desalination.  Here we report the study of
    % tailoring the ion transport through atomically-thin porous
    % graphene with pore sizes of 20$\pm$10 nanometers, and a pore
    % density of 1.25$\pm$0.25$\times$10$^{10}$ cm$^{-2}$. By
    % electrostatically gating a sheet of porous graphene, we
    % demonstrate a considerable reduction of salt flux by up to 55\%
    % and 85\% using a gate voltage, based on our experiments and
    % theoretical predictions, respectively. We systematically
    % investigate the effects of salt concentration and species,
    % including the development of a theory to model the electrolyte
    % diffusion through a nanopore in a sheet of gated graphene. The
    % interplay between graphene quantum capacitance and electrical
    % double layer is identified to be the main mechanism responsible
    % for the observed salt flux reduction, when the pore size is
    % comparable to the Debye length. Our findings reveal a new degree
    % of freedom controlling the membrane potential and electrolyte
    % permeation through porous two-dimensional materials.
  }
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}
\linenumbers
\section{Introduction}
\label{sec:intro}
A central objective in the emerging field of nanofluidics is to
control transport phenomena at nanometer scale to explore new
technological opportunities. The reduction of characteristic
dimensions in nanofluidics gives rise to substantially different
transport properties in comparison to their bulk counterparts
\cite{Schoch_2008}. In nanopores, for example, the effect of
surface-mediated transport becomes dominant when the pore size is
comparable to the characteristic length of diffusion, enabling new
applications such as ionic diodes \cite{Karnik_2007}, field-effect
transistors \cite{Nam_2009}, desalination \cite{Heiranian_2015}, and
nanopore-based DNA sequencing \cite{Heerema_2016,Garaj_2013}. Porous,
atomically thin two-dimensional (2D) materials, such as graphene, are
emerging as ideal membrane materials due to the ultimate permeation
\cite{Suk_2010,Jiang_2009,Celebi_2014,Koenig_2012,Drahushuk_2012}. Early
research has suggested that using a sheet of nanoporous graphene can
act as a desalination membrane by having the pore sizes smaller than
the hydrated radii ($\sim$1 nm), thereby hindering ion passage without
losing water permeability
\cite{Cohen_Tanugi_2012,Suk_2014,Cohen_Tanugi_2014,Cohen_Tanugi_2015,O_Hern_2014,O_Hern_2015,Surwade_2015,Walker_2017,Ghosh_2018}. Nevertheless,
in practice, the fabrication of subnanometer porous 2D membranes over
a large area with atomic precision remains technically challenging
\cite{Suk_2014,Rollings_2016,O_Hern_2012,Wang_2017}. Moreover, a small
portion of excessively large pores or tears can significantly diminish
salt rejection.  Very recently, a number of reports have explored the
effect of surface charges on the selectivity of ion transport through
graphene nanopores, as a consequence of the electrostatic interactions
between solvated ions and the charged functional groups attached on
graphene, indicating that even large pores ($>$ 30nm) can be ion
selective \cite{Rollings_2016,Surwade_2014}. However, it remains
controversial that if the surface charges can result in a degree of
salt rejection, as the surface charge and potential cannot be
precisely quantified and are sensitive to the surroundings as well as
the history of chemical treatment \cite{Li_2008}. There remains a lack
of fundamental understanding of the interactions between the charged
2D nanopores and the electrolyte solution. In this report, we
investigate electrolyte permeation through graphene nanopores that are
significantly larger than the hydrated radii. We model the surface
potential by taking into account of the elementary electronic
properties of graphene and demonstrate that a high degree of salt
rejection can be achieved by electrostatic gating.

\section{Results and discussions}
\label{sec:res}

\subsection{Free diffusion through porous graphene}
\label{sec:res-1}

\Fig{} \ref{fig:1}a depicts the experimental setup characterizing
ionic diffusion through a sheet of double layer porous graphene
(PG). Two reservoirs, namely the high-concentration reservoir (HCR)
containing electrolyte solution with molar concentration $c_0$, and
the low-concentration reservoir (LCR) containing de-ionized (DI)
water, are separated by a PG membrane supported by a polycarbonate
track-etched (PCTE) film. Magnetic stirrers are used to minimize the
effects of external concentration polarization, and a conductivity
probe is placed in the LCR to monitor the ionic concentration as a
function of time. The membrane is attached to a piece of copper tape
connected to a voltage source applying an electrical bias
$V_{\mathrm{G}}$, with the LCR grounded (\Fig{} \ref{fig:1}a
inset). We fabricated the PCTE-supported PG as schematically shown in
\Fig{} \ref{fig:1}b, using the protocol developed by our group
\cite{Choi_2018} (details see Supplementary Section
\ref{SI-sec:exp}). The normally distributed, circular pores with pore
diameters of 20$\pm$10 nm and a pore density of
1.25$\pm$0.25$\times$10$^{10}$ cm$^{-2}$ (\Fig{} \ref{fig:1}c inset)
over a large area were perforated on graphene.  By optimizing the
subsequent transfer process, a high surface coverage ($>$ 98\%) of PG
on PCTE was obtained (\Fig{} \ref{fig:1}c), allowing reliable
characterization of the ion transport behavior (details see
Supplementary Section \ref{SI-sec:exp}).  The system presented here
allows us to systematically investigate the diffusive flux of ions
across a sheet of PG, $\boldsymbol{J}$, driven by the gradient of
electrochemical potential in the electrolyte medium.

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity increases linearly with
time in the LCR (e.g. \Fig{} \ref{fig:2}a) within the measurement
duration (1 h), suggesting the concentration gradient across the
membrane remains constant. The diffusive fluxes (blue bars in \Fig{}
\ref{fig:2}b) of different salts were obtained by converting the
conductivity-time profiles using calibration curves (Supplementary
Section \ref{SI-sec:measure}). The theoretical diffusive fluxes
through PCTE are obtained based on the assumption that the membrane
consists of cylindrical pores with uniform diameter and a low
tortuosity $\tau$ of $\sim{}$1.2 \cite{O_Hern_2012}:
\begin{equation}
  \label{eq:j-pcte}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi r_{\mathrm{PCTE}}^{2} N_{\mathrm{p}} D_{\mathrm{\pm}} \Delta c}{L \tau}
\end{equation}
where $r_{\mathrm{PCTE}}$ is the PCTE pore radius, $N_{\mathrm{p}}$ is
numbers of pore per area, $D_{\mathrm{\pm}}$ is the salt diffusivity
(average of cation and anion diffusivity), $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$ and $L$ is the thickness of the PCTE membrane.
By using the salt diffusivity values (Supplementary Table
\ref{SI-tab:diff}), the calculated diffusive fluxes (red bars in
\Fig{} \ref{fig:2}b) show reasonable agreement with the experimental
values.

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
with those of bare PCTE membrane, suggesting that the transport
resistance of PCTE $R_{\mathrm{PCTE}}$, is dominant over that of PG
without gating, $R_{\mathrm{G}}^{0}$. Using a simple circuit model, we
estimate the ratio $\delta$ between $R_{\mathrm{G}}^{0}$ and
$R_{\mathrm{PCTE}}$, i.e.,
$\delta = R_{\mathrm{PCTE}}/R_{\mathrm{G}}^{0}$, to be $\sim$2.2,
which is in good agreement with the value derived from the
experimental data in \Fig{} \ref{fig:2}b (details see Supplementary
Section \ref{SI-sec:R-model}).
% \todo[inline]{Make sure that the
  % values used are correct.}

% The diffusive transport resistance through graphene pores scales with $\sim$ 1/D while
% for a PCTE pore, the scaling follows $\sim$ L/D$^2$ \cite{O_Hern_2014},
 % D being the respective diameter and L the channel length. 
% Taking into account that there are 10 graphene pores
% per PCTE pore on average, the transport resistance ratio of R$_{graphene}$/R$_{PCTE}$ is roughly 1/30,
 % indicating that the PCTE dominates the overall ion transport, which is in line 
 % with the observed behavior.

%Interestingly, we observed a pronounced decrease of diffusive flux in KCl,
%K$_{2}$SO$_{4}$ and K$_{3}$[Fe(CN)$_{6}$] systems but not for the rest
%of the salts. We hypothesized that the strong cation-$\pi$ interactions
%between K$^{+}$ cations and graphene slow down the migration of cations
%along the surface-mediated transport pathway \cite{Sun_2014}.

\subsection{Salt rejection induced by electrostatic gating}
\label{sec:res-2}

We next discuss the effects of electrostatic gating on graphene. In
order to avoid any interference from the conductivity probe to the
membrane potential, the ``three-interval'' method, which has been used
in controlling the membrane potential in mesoporous carbon membranes
\cite{Surwade_2014}, is adopted here. \Fig{} \ref{fig:2}a
illustratively represents how the measured conductivity evolves with
time. Specifically, in the first interval, the gate voltage source is
turned off and the conductivity probe is on, allowing us to obtain the
conductivity-time profile with an average slope $s_{1}$ that
corresponds to the diffusive flux across the PG-PCTE membrane. In the
second interval, a voltage $V_{\mathrm{G}}$ is applied to the membrane
after turning off the conductivity probe, followed by the third
interval $s_{3}$, in which the conductivity probe is turned on again
to give the corresponding slope. A control experiment was performed to
ensure the copper tape was not in direct contact with the ionic
solution (Supplementary Section \ref{SI-sec:copper}). The slope
corresponding to the second interval $s_{2}$, as the conductivity
probe is off, is determined by linear interpolation of the
conductivities from the endpoint of interval 1 to the starting point
of interval 3. A decrease of slope observed in interval 2 indicates
that the total flux through PG-PCTE is reduced upon gating (i.e. salt
penetration through the PG pores is hindered). We define such
experimentally measured reduction rate as $\eta$, which is equivalent
to the decrease of total diffusive flux through PG-PCTE as follows:
\begin{equation}
  \label{eq:rejection}
  \eta = \frac{\bar{s} - s_{2}}{\bar{s}} = \frac{\boldsymbol{J}_{\mathrm{PG0}}
    - \boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})}{\boldsymbol{J}_{\mathrm{PG0}}}
\end{equation}
where $\bar{s}$ is the average slope of $s_{1}$ and $s_{3}$, namely
$(s_{1} + s_{3})/2$, and
$\boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})$ is the diffusive flux
across the PTCE-supported PG double layer as a function of
$V_{\mathrm{G}}$. Since graphene and PCTE membranes effectively form a
transport system with series diffusive resistance $R_{\mathrm{G}}$
(tunable by $V_{\mathrm{G}}$) and $R_{\mathrm{PCTE}}$ (independent of
$V_{\mathrm{G}}$), respectively, we define the effective degree of
salt rejection from bare PG $\xi$ as a function of $V_{\mathrm{G}}$ (
details see Supplementary Section \ref{SI-sec:R-model}) as:
\begin{equation}
\label{eq:xi-def}
\xi = 1 - \frac{R_{\mathrm{G}}(V_{\mathrm{G}}=0)}{R_{\mathrm{G}}(V_{\mathrm{G}})} = \frac{(\delta+1) \eta}{\delta \eta + 1}
\end{equation}
The factor $\xi$ is a reliable and stable measure of salt rejection
through the PG membrane itself upon gating, as the salt species and
initial conditions may induce a degree of measurement uncertainty
between different samples. Note that the $\xi=1$ limit represents
perfect salt rejection in which
$R_{\mathrm{G}}(V_{\mathrm{G}}) \to \infty$. On the contrary $\xi=0$
indicates $R_{\mathrm{G}}$ remains unchanged with $V_{\mathrm{G}}$.

Using 0.1 mM KCl solution in HCR, we obtained $\xi$ as a function of
$V_{\mathrm{G}}$ within $\pm1.25$ V, before triggering the
electrochemical reactions (\Fig{} \ref{fig:2}c). We observe an
asymmetric dependence of $\xi$ with respect to $V_{\mathrm{G}}$, with
a higher degree of salt rejection for $V_{\mathrm{G}}>0$. where
positive carriers (holes) are induced in graphene. We notice that an
increase of $\xi$, or a decrease of diffusive flux through PG with
$V_{\mathrm{G}}$, showing an inverse trend compared with those
observed in ionic field effect transistors (IFETs) in which the
diffusive flux increases with the gate voltage
\cite{Nam_2009,Cheng_2018}. A further increase of the KCl
concentration $c_{0}$ influences the attainable degree of salt
rejection. $\xi$ values measured at $V_{\mathrm{G}}$ = 1.25 V,
$\xi_{\mathrm{max}}$, as a function of $c_{0}$ from 10$^{-4}$ to
10$^{-2}$ M KCl (see Supplementary Section \ref{SI-sec:debye}) exhibit
a power law dependency (\Fig{} \ref{fig:2}d). The
$\xi_{\mathrm{max}}-c_{0}$ relation can be nicely fitted by a power
law function following
$\xi_{\mathrm{max}} c_{0}^{0.52} = \mathrm{constant}$. As the Debye
length in solution, $\lambda_{\mathrm{D}}$, is inversely proportional
to $c^{0.5}$, we infer that the $V_{\mathrm{G}}$-induced salt
rejection originates from the modulation of the electrical double
layer (EDL), as will be discussed later.

\subsection{Self-consistent ion transport theory}
\label{sec:theory}

In order to quantitatively understand the observed $V_{\mathrm{G}}$
dependence, we develop a theory to model the coupling of graphene's
elementary transport properties and EDL, in order to quantify the
ionic transport through a graphene nanopore. Under the assumption that
the time scale for the bulk concentration change is significantly
longer than that of ionic diffusion, the pseudo-steady state
approximation holds. Accordingly, the steady-state Nernst-Planck
equation describing the ionic transport in an electrolyte solution is
given by \cite{MacGillivray_1968}:
\begin{equation}
  \label{eq:pnp}
  \nabla \cdot \boldsymbol{J}_{i} = -\nabla \cdot (\frac{D_{i}}{k_{\mathrm{B}}T} c_{i} N_{\mathrm{A}} \nabla \mu_{i}) = 0
\end{equation}
where $\boldsymbol{J}$ is the mass flux, subscript \textit{i}
corresponds to the ionic species (anion or cation), $D$ is the diffusivity,
$k_{\mathrm{B}}$ is the Boltzmann constant, $T$ is temperature, $c$ is
the molar concentration, $N_{\mathrm{A}}$ is the Avogadro constant,
and $\mu_{i}$ is the electrochemical potential. Under the assumption
of ideal solution, it follows \cite{Kilic_2007}:
\begin{equation}
  \label{eq:mu}
  \nabla \mu_{i} = k_{\mathrm{B}} T \nabla \ln x_{i} + z_{i} e \nabla \psi
\end{equation}
 $x$ is the molar fraction in solution, $z$ is the ionic valence,
$e$ is the unit charge, and $\psi$ is the electric potential. On the
other hand, the Poisson equation describing the electric potential
distribution is given by:
\begin{equation}
  \label{eq:poisson}
  \nabla \cdot (\varepsilon_{\mathrm{m}} \varepsilon_{0} \nabla \psi)
  =
  - N_{\mathrm{A}} e \sum_{i} c_{i} z_{i}
\end{equation}
where $\varepsilon_{\mathrm{m}}$ is the relative permittivity of
individual materials in the system, including water and the internal
Stern layer\cite{Tian_2017}, and
$\varepsilon_{0}$ is the vacuum permittivity. When applying
$V_{\mathrm{G}}$ to graphene adjacent to the electrolyte solution,
charges (electrons or holes) are induced in graphene; the
electroneutrality of the entire system suggests:
\begin{equation}
  \label{eq:electro-neutral}
  \sigma_{\mathrm{G}} S_{\mathrm{G}} + \sum_{i} \int_{\Omega} z_{i} c_{i} N_{\mathrm{A}} e \mathrm{d}^{3} \Omega= 0
\end{equation}
where $\sigma_{\mathrm{G}}$ is the charge density in graphene,
$S_{\mathrm{G}}$ is the total area of graphene, and $\Omega$
corresponds to the entire volumetric domain of electrolyte
solution. Note that the graphene surface potential,
$\psi_{\mathrm{G}}$, is not equivalent to $V_{\mathrm{G}}$ due to
change of graphene's work function upon charging, known as the
graphene quantum capacitance effect\cite{Xia_2009} (\Fig{}
\ref{fig:3}a) following:
\begin{equation}
  \label{eq:Vg}
  V_{\mathrm{G}} = \Delta \phi_{\mathrm{G}} + \psi_{\mathrm{G}}
\end{equation}
where
$\Delta \phi_{\mathrm{G}} = \phi_{\mathrm{G}}(V_{\mathrm{g}}) -
\phi_{\mathrm{G}}(V_{\mathrm{g}}=0)$ is the change of graphene’s work
function. The properties of the PG samples fabricated is close to
intrinsic graphene, with charge density at $V_{\mathrm{G}}=0$ measured
to be 1.4$\times$10$^{11}$ \textit{e}$\cdot$cm$^{-2}$, corresponding
to a work function difference of 48 meV from graphene's Dirac point
(details see Supplementary Section \ref{SI-sec:charge-dens}), much
smaller than the $\Delta \phi_{\mathrm{G}}$ induced by gating.
Therefore we adapt the simplification that the charge neutrality point
of graphene coincides with the Dirac point at $V_{\mathrm{G}}$ =
0. The elementary electronic properties of graphene give:
\begin{equation}
  \label{eq:delta-phiG}
  \Delta \phi_{\mathrm{G}} = \mathrm{sign}(\sigma_{\mathrm{G}}) \frac{\hbar v_{\mathrm{F}}}{e}
  \sqrt{\frac{\pi |\sigma_{\mathrm{G}}|}{e}}
\end{equation}
where $\hbar$ is the reduced Planck constant, and
$v_{\mathrm{F}}=1.1\times10^{6}$ m$\cdot$s$^{-1}$ is the Fermi
velocity of graphene (more details find Supplementary Section
\ref{SI-sec:charge-dens}).  Note that the PG film fabricated
experimentally is double-layer turbostratic (randomly aligned)
graphene, in which we assume the
$\Delta \phi_{\mathrm{G}} - \sigma_{\mathrm{G}}$ dependence follows
Eq. \eqref{eq:delta-phiG}. The theory proposed here was solved
self-consistently by discretizing
Eqs. \eqref{eq:pnp}-\eqref{eq:poisson} using the finite-element method
(FEM), in which the graphene surface potential is coupled with
Eqs. \eqref{eq:electro-neutral}-\eqref{eq:delta-phiG} that were solved
simultaneously to reach convergent numerical solutions of
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ for a given
$V_{\mathrm{G}}$. Clearly, the above model yields symmetric
characteristics for $\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ with
respect to $V_{\mathrm{G}}$ due to the symmetric band structure of
graphene (Eq. \eqref{eq:delta-phiG}). However, for graphene
synthesized by chemical vapor deposition (CVD) used in our
experiments, it is well-recognized that the electron traps are
inherently introduced during synthesis and patterning
\cite{Dean_2010}, effectively reducing the charge density and surface
potential on graphene $V_{\mathrm{G}}<0$ (Supplementary \Fig{}
\ref{SI-fig:trap}), which is related to the fact that observed $\xi$
is much lower within the $V_{\mathrm{G}}<0$ regime in \Fig{}
\ref{fig:2}c. With the nonideality in mind, hereafter, we compare the
calculations and experiments in the regime of $V_{\mathrm{G}}>0$. For
example, \Fig{} \ref{fig:3}c presents the calculated
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ as a function of
$V_{\mathrm{G}}$ considering the ionic diffusion through a single 20
nm-diameter nanopore drilled on a sheet of semi-infinite double layer
graphene that separates HCR containing KCl solution at $c_{0}$ = 0.1
mM, and LCR at $c_{0}/10$ in axisymmetric coordinates (details see
Supplementary Section \ref{SI-fig:scheme}). Indeed, the interplay
between graphene quantum capacitance and the EDL significantly reduces
the graphene surface potential $\psi_{\mathrm{G}}$ to $\sim$0.3 V at
$V_{\mathrm{G}}$ = 1.25 V, corresponding a surface charge density
$\sigma_{\mathrm{G}}$ of $\sim$0.08 C$\cdot$m$^{-2}$. Note that as we
mentioned earlier, an important merit of the 2D nanopore system
considered here is that the surface charge density can be precisely
determined and controlled, rather than being treated as a fitting
parameter as in most of the 2D nanopore literature
(e.g. Ref. \citenum{Rollings_2016}). In addition, we anticipate that
the gate-controlled ionic transport can also be achieved by replacing
graphene with other nanoporous conducting 2D materials, although the
operational gate voltage window required to achieve the same degree of
salt rejection may be narrower, because graphene has the smallest
quantum capacitance (near the Dirac point) among all known 2D
materials \cite{Tian_2016}.

\subsection{Salt rejection mechanism}
\label{sec:mechanism}

The numerical procedure described above allows us to resolve the
concentration and electric potential profiles near a graphene
nanopore for a given $V_{\mathrm{G}}$. Since the ionic flux is driven
by both $\psi$ and $c$ fields (see Eq. \eqref{eq:pnp} and \eqref{eq:mu}),
we focus on the electrochemical potential $\mu_{i}$, which represents
the combined driving force, to reveal its dependence on the applied
$V_{\mathrm{G}}$. Following the same system considered in \Fig{}
\ref{fig:3}b, the calculated axisymmetric electric potential $\psi$
and the relative electrochemical potentials,
$\Delta \mu_{\mathrm{K^{+}}}$ and $\Delta \mu_{\mathrm{Cl^{-}}}$, for
$V_{\mathrm{G}}$ = 0.75 V are shown in \Fig{} \ref{fig:4}a and
\ref{fig:4}b, respectively. The reference point of the electrochemical
potentials is set at the bulk phase in the HCR. As expected, the
electric potential reaches the maximum at the graphene surface (with
$\psi_{\mathrm{G}}$ of $\sim$100 mV) and decays towards the bulk
solution phase, forming an EDL surrounding the graphene surface. Since
the Debye length $\lambda_{\mathrm{D}}$ is comparable to the
pore radius $r_{\mathrm{G}}$, the electric potential at the nanopore
center remains at $\sim$25 mV, comparable to the thermal energy at
room temperature ($k_{\mathrm{B}}T$ = 26 meV). Consequently, it is
evident that the potential barrier is sufficient to modulate the
diffusive flux.

We further reveal the cation and anion transport pathways by looking
into their electrochemical potentials (\Fig{} \ref{fig:4}b). By
increasing $V_{\mathrm{G}}$, a more positive $\psi_{\mathrm{G}}$ on
graphene surface reduces the cation concentration at the pore edge due
to the electrostatic interactions, thereby increasing its
concentration gradient at the pore center. On the other hand, the
anion concentration near the pore edge increases exponentially, such
that the concentration gradient at the pore center becomes small (see
Supplementary \Fig{} \ref{SI-fig:conc}). We also observe that
$\Delta \mu_{+}$ is dominated by the concentration, while
$\Delta \mu_{-}$ is much less than $\Delta \mu_{+}$ due to the balance
between the diffusion and drift potentials (see Supplementary \Fig{}
\ref{SI-fig:potential}). The observations confirm the distinct ionic
transport pathways upon applying a positive $V_{\mathrm{G}}$: pore
center for cations and pore edge for anions. \Fig{} \ref{fig:4}c
presents the calculated \textit{z}-component of the cation and anion
fluxes through the pore at the graphene plane (z = 0),
$\boldsymbol{J}_{z}$ ,as a function of the normalized radius,
$r/r_{\mathrm{G}}$ , where $r$ is the radial coordinate and
$r_{\mathrm{G}}$ is the pore radius, at different $V_{\mathrm{G}}$
values. Note that $\boldsymbol{J}_{z}$ is negative because the LCR is
placed at z$<$0 in the simulation box. At $V_{\mathrm{G}}$ = 0 (pure
diffusion), the anion and cation flux profiles are identical, with a
higher flux near the pore edge, which is expected, as the
concentration gradient is higher. By gradually increasing
$V_{\mathrm{G}}$, both fluxes are reduced throughout the pore, while
the degree of reduction is more pronounced at the pore edge for the
cation, and at the pore center for the anion, following the scenarios
we discussed above. Accordingly, the integrated fluxes across the
pore, $|\boldsymbol{J}_{z}|$, are reduced with $V_{\mathrm{G}}$
(\Fig{} \ref{fig:4}d). A total flux reduction of $\sim$90\% is
predicted. Another interesting finding here is that, by increasing
$V_{\mathrm{G}}$, the nanopore transport preferentially allows cations
over anions, or in other words, the anion flux is more reduced with
$V_{\mathrm{G}}$ (see \Fig{} \ref{fig:4}c), known as the ion
selectivity of graphene nanopore \cite{Rollings_2016}. In our setup,
the imbalance between cation and anion fluxes near the graphene
nanopore is compensated by the EDL formation at the counter-electrode
in LCR, which maintains the electroneutrality of the system. Notably,
since atomically thin graphene samples are used, the mechanism of salt
rejection is clearly distinct from that in capacitive deionization
\cite{Biesheuvel_2010}, where electrodes with large surface area are
required.

The above mechanistic findings, nevertheless, do not fully clarify the
experimentally observed reduction of diffusive flux upon
gating. Indeed, the same physical mechanism also governs the ionic
transport through an IFET, in which a nonzero electric potential at
the nanopore center usually results in an increase of ionic
conductivity \cite{Nam_2009,Lee_2015,Feng_2016}. To this end, we
further increased the graphene surface potential $\psi_{\mathrm{G}}$
considering the same system (details see Supplementary Section
\ref{SI-sec:simu-res}). Intriguingly, by increasing
$\psi_{\mathrm{G}}$ larger than 400 mV, the ionic flux starts to
increase, reversing the trend observed at the low $\psi_{\mathrm{G}}$
regime, due to (i) the drift flux become dominant over diffusion flux
and (ii) the concentration near the pore edge (preferred path) greatly
increases (details see Supplementary Section \ref{SI-sec:simu-res},
Supplementary Figs. \ref{SI-fig:reverse} and
\ref{SI-fig:large-V}). This level of $\psi_{\mathrm{G}}$ cannot be
reached experimentally in graphene membrane, as this requires a
$V_{\mathrm{G}}$ larger than 2.0 V, which triggers the electrochemical
reactions \cite{Toh_2011}. On the other hand, in an IFET, the electric
potential on the pore wall is considerably higher, equivalent to
$V_{\mathrm{G}}$ due to the metallic nature of gate electrode (having
an infinitely-large quantum capacitance). We therefore conclude that
the quantum capacitance-induced nonlinear damping in the surface
potential results in the observed salt rejection.  Following the above
discussions, we further investigate the physical limits for the biased
graphene nanopores. Specifically, the salt rejection characteristics
are controlled by: (i) the overlap of EDL inside the nanopore,
essentially controlled by two length scales of $\lambda_{\mathrm{D}}$
and $r_{\mathrm{G}}$, and (ii) the graphene surface potential
$\psi_{\mathrm{G}}$ controlled by $V_{\mathrm{G}}$ following
Eq. \eqref{eq:Vg}. To this end, we calculate $\psi_{\mathrm{G}}$ as a
function of $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:5}a). Clearly, when the bulk
concentration $c_{0}$ increases, a higher $V_{\mathrm{G}}$ is required
to reach the same level of $\psi_{\mathrm{G}}$. Consequently, as
revealed in \Fig{} \ref{fig:5}b, the calculated salt rejection factor
$\xi$ increases with both $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$, in line with the experimentally observed
$\xi - c_{0}$ dependence. We also find the overall trends for
$\psi_{\mathrm{G}}$ and $\xi$ with respect to
$\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ and $V_{\mathrm{G}}$ are very
similar. More importantly, based on our theoretical prediction, with
$V_{\mathrm{G}}$ = 1.25 V and $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$
= 1.0, a high value of $\xi$ up to $\sim$0.95 can be achieved.  We
notice that although the above analysis suggests a salt rejection up
to 1 may be achived by further reducing the pore size (i.e. increasing
the $\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ ratio), our theory, derived
from on the macroscopic transport equations by treating ions as
zero-volume charges, may not be valid when $R_{\mathrm{G}}$< 5
nm\cite{Jain_2015}. In addition, the specific ion-graphene
interactions \cite{Rollings_2016} are not taken into account. Advanced
theoretical frameworks that bridge continuum and molecular models are
required to elucidate electrolyte transport approaching the small pore
limit.
% Nevertheless
% the salt rejection of the gated graphene platform may be further
% improved after considering the following aspects: (i) the mean-field
% theory used here may be incorrect to be applied to pores smaller than
% 5 nm, where individual particle properties of solvated ions can be
% important \cite{Jain_2015}, (ii) the specific ion-graphene adsorption
% may also be included to account for the pore selectivity between salt
% species \cite{Rollings_2016}.

\subsection{Effects of salt species}
\label{sec:salts}

Finally, we examine other salt species and compare the degree of salt
rejection between experiments and simulations. \Fig{} \ref{fig:6}a
presents the measured salt rejection $\xi$ as a function of
$V_{\mathrm{G}}$ at $c_{0}$ = 0.1 mM, for the seven salt species
considered here, including NaCl, LiCl, KCl, MgSO$_{4}$, CaCl$_{2}$,
K$_{2}$SO$_{4}$, and K$_{3}$[Fe(CN)$_{6}$] (details see Supplementary
Section \ref{SI-sec:simu-res}). Similar to the KCl system discussed
earlier, the salt rejection is more pronounced in the positive
$V_{\mathrm{G}}$ regime. The degree of salt rejection deceases with
the salt valance; for example, the measured $\xi$ values at
$V_{\mathrm{G}}$ = 1.25 V decreases from $>$ 0.5 for monovalent NaCl
to -0.04 for multivalent K$_{3}$[Fe(CN)$_{6}$]. This observation is
endorsed by our simulations, in which we calculate $\xi$ as a function
of $V_{\mathrm{g}}$ using the same simulation setups (\Fig{}
\ref{fig:6}b). The calculated $\xi_{\mathrm{max}}$ values are
quantitatively compared with experiments (\Fig{} \ref{fig:6}c), and
the same trend is observed. Indeed, because the Debye length
$\lambda_{\mathrm{D}}$ decreases with the salt valence, the graphene
surface potential appears to decrease in the multivalent salt systems,
thereby reducing $\xi$ (see \Fig{} \ref{fig:5}b). The calculated and
experimental $\xi_{\mathrm{max}}$ as a function of
$\lambda_{\mathrm{D}}$ are consequently plotted, for the seven salt
species considered here (\Fig{} \ref{fig:6}d). Similar to the
experimental observations, small difference between the $\xi$ of salts
with same valence (e.g. LiCl, NaCl and KCl) is also predicted from the
simulations due to the variation in ionic mobilities. We notice that
without any fitting parameters, our theory can predict the
experimental values reasonably well.  In addition, the
$\xi_{\mathrm{max}}-\lambda_{\mathrm{D}}$ dependence is approximately
linear, suggesting that the effect of Debye length is dominant over
the salt diffusivity. This finding also explains the concentration
dependence of $\xi_{\mathrm{max}}$ observed in \Fig{} \ref{fig:2}d, as
$\lambda_{\mathrm{D}} \propto c_{0}^{-0.5}$. Further including the
specific interaction of ionic species with the graphene membrane
\cite{Ghosh_2018} may give even better agreement with experimental
results.

In summary, we have presented a comparative experimental and modelling
study on macroscopic salt rejection through a sheet of large-area
porous graphene under electrostatic gating. We show that due to a
small quantum capacitance of atomically-thin graphene, the graphene
surface potential is considerably lower than the applied gate
voltage. As a result, the subtle redistribution of the electrochemical
potentials create new pathways for ionic transport, which in turn lead
to a considerable degree of salt rejection. We report a high degree of
salt rejection at $V_{\mathrm{G}}$ = 1.25 V by approximately one order
of magnitude, based on our experiments and theoretical predictions,
respectively. We demonstrate that the observed salt rejection
positively correlates with the Debye length that nonlinearly modulates
the graphene surface potential and charge density. The experimental
results and fundamental principles presented here open an avenue
towards realization of atomically thin 2D porous membranes for
advanced separation process in electrolyte solutions.

\section{Methods}
\label{sec:methods}

\textbf{Graphene synthesis}: Graphene was grown in a cold-wall,
low-pressure chemical vapor deposition system (BM 4 inch, Aixtron SE)
on surface-pretreated copper foil (Alfa Aesar No. 46986). Synthesis was
carried out at 950 $^{\circ}$C for 3 min using ethylene after 30 min of
hydrogen/argon annealing. A first graphene layer was subsequently
transferred to a second one to yield double layer graphene using a
PMMA transfer method. A glass slide was used as the support for
subsequent patterning of the graphene, PMMA was removed using acetone
ultrasonication and subsequent isopropanol (IPA) rinsing. The entire
process is described in detail in Ref. \citenum{Choi_2018}.

\vspace{1em}
\noindent
\textbf{Graphene patterning and transfer to PCTE}: Spherical block
copolymer (s-BCP) based patterning has enabled the manufacturing of
perforated graphene membranes having pores in the range of 10 nm to 50
nm on average up to wafer-scale\cite{Choi_2018}. In brief, s-BCP
(polystyrene-block-polymethylmethacrylate, PS-b-PMMA, 195-b-20,
Polymer Source Inc, Canada, 1 weight\% in anhydrous toluene) is
spin-coated from solution and thermally annealed at 220 $^{\circ}$C
for 6 h in vacuum. Microphase separation of s-BCP leads to distinct
phases: the minor phase will form spheres (PMMA) and the major phase
will form the matrix (PS). Oxygen plasma (8 , 100 W, 25 mTorr, Oxford
Instruments plc) removes the top layer PS from the PMMA spheres, which
are subsequently etched in glacial acetic acid to form a porous
etch mask directly on top of the double layer graphene. Anisotropic
etching using oxygen beam milling (10 s, 100 mA, 600 V acceleration
voltage, Oxford IonFab 600) leads to perforation of the underlying
graphene through the porous PS mask. Polymer residues are removed by
thermal annealing in hydrogen:argon (9:1) at 400 $^{\circ}$C. The
patterned graphene is transferred to PCTE membranes (hydrophobic, 400
nm pores, from Sterlitech) by baking the PCTE briefly to the
perforated graphene at 180 $^{\circ}$C followed by a lift-off etch in
buffered hydrofluoric acid (BHF) as described by\cite{Choi_2018}. A
schematic of the process is shown in \Fig{} S1.

\vspace{1em}
\noindent
\textbf{Membrane preparation}: The ion-diffusion and gating experiments are
performed in a custom-made diffusion cell. An initial dip-wetting
procedure using ethanol/water 1:1 is performed to wet the membrane,
followed by extensive washing and inserting the membrane. The
diffusion cell is filled immediately after mounting the membrane with
DI water to prevent drying. The ion solution is placed on the HCR
side, distilled water on the LCR side, with a volume of 7.33 mL for
each cell. Bar-stirrers close to the membrane ensure well-mixed
solutions avoiding potential external concentration polarization
effects. Water channels in the fixture ensure stable measurement
temperatures using a chiller. The temperature for all measurements is
kept at 25 $^{\circ}$C. A conductivity probe (eDAQ, Pty Ltd.,
Australia), inserted on the LCR side measures the conductivity, which
is recorded with time steps of 2 s. The membrane is kept wet for the
entire series of experiments, multiple rinsing steps with DI water are
performed after each experiment and change to a different salt
solution. The salt solutions were prepared immediately before the
experiments using MilliQ water that was additionally distilled to
increase purity level. The distilled water, has a conductivity of
0.0012 mS/cm with a pH of 6.3, no buffer has been used in these
experiments. 7 salts were used: KCl, NaCl, LiCl, K$_{2}$SO$_{4}$,
CaCl$_{2}$, MgSO$_{4}$ and K$_{3}$[Fe(CN)$_{6}$]. The salts were
purchased from Sigma Aldrich and used without further purification,
salt solutions were prepared by subsequent dilution of a base solution
(1 M) to the final concentrations.

\vspace{1em}
\noindent
\textbf{Ion diffusion experiments}: Baseline measurements were
performed using bare PCTE (b-PCTE) by extracting the diffusion rates
of 3 measurements and averaging the resulting values (see
Supplementary Section \ref{SI-sec:measure}). For the measurement, DI water was filled on
the LCR side, where a 0.1 mM salt solution is placed on the HCR side
of the fixture. The conductivity increase was recorded over 1 h before
rinsing and replacing both sides with fresh solutions.

\vspace{1em}
\noindent
\textbf{Ion gating experiments}: The salt rejection behavior of the graphene
was characterized using patterned graphene on PCTE (PG-PCTE) in a
3-interval-type measurement based on salt diffusion: 1h per interval
with (i) no bias (ii) bias and (iii) no bias, ruling out possible
electrochemical effects at the membrane that might mislead
interpretation of the conductivity reading. The voltage is applied via
copper tape so that the membrane acts as a working electrode (WE). A
platinum wire (length 3 cm, diameter 0.6 mm) placed in the low
concentration side of the diffusion cell with a distance of $\sim$3.5 cm
from the PG-PCTE membrane is counter- (CE) and reference electrode
(RE) simultaneously, forming effectively a two-electrode setup. An
electrochemical workstation (PGSTAT204, Metrohm Autolab S.V.) is used
to apply the selected voltage.

\vspace{1em}
\noindent
\textbf{Numerical Simulations}: Numerical simulations using finite
element method (FEM) were carried out using COMSOL Multiphysics
5.3a. A single pore in graphene is simulated, with the horizontal
length of the simulation box set to 20 times of the nanopore size. The
transport of ionic species was solved by the Nernst-Planck equation,
with the electrostatic potential further given by the Poisson
equation. Since the surface on graphene cannot be predetermined, we
performed all calculations by setting the values of
$\psi_{\mathrm{G}}$. The surface charge of graphene is further
extracted and used to reconstruct the value of actual $V_{\mathrm{G}}$
applied to graphene. We examined the convergence of the solutions by
performing simulations with varied mesh sizes. For further details
about the simulation setup please refer to Supplementary Section
\ref{SI-sec:numer}.

\begin{acknowledgement}
  R.M.W. acknowledges Dr. Kyoungjun Choi from ETH Zurich for the measurement of the electron mobility 
  of patterned graphene and is grateful for the support from the Binnig and Rohrer
  Nanotechnology Center of ETH Zurich and IBM Zurich. T.T. and
  C.J.S. are grateful for the financial support from the ETH startup
  funding. 
\end{acknowledgement}

\section{Author Contributions}
\label{sec:author}

H.G.P and C.J.S conceived the experiments. R.M.W and K.Y. measured the ion
diffusion and gating experiments. T.T performed numerical
simulations. All authors contributed to writing and discussion.

\section*{}
\label{sec:ref}
\bibliography{ref}

\clearpage
\section{Figures}
\label{sec:figs}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/fig1.pdf}
  \caption{\textbf{Experimental setup and nanopore characterization}.
    \textbf{a}. 3D schematic diagram showing the experimental
    setup. Inset: photograph of the PG-PCTE membrane supported by
    Kapton and copper tapes. (Scale bar: 1 cm.) \textbf{b}. 3D schematic
    of the nanoscale cross-sectional structure of the PG-PCTE
    membrane. \textbf{c}. SEM image of the PG-PCTE membrane, showing an
    excellent coverage of graphene over the PCTE substrate, scale bar:
    20 $\mathrm{\mu}$m.  Upper right inset: high-resolution SEM image
    of individual nanopores on PG fabricated by the patterning
    technique. (Scale bar: 500 nm.) Bottom right inset: histogram of the
    pore diameter distribution revealed by the SEM
    image. \textbf{d}. Schematic of the nanoscale ionic transport
    through a gated graphene nanopore.}
  \label{fig:1}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/fig2.pdf}
  \caption{\textbf{The experimental evidence of salt rejection.}
    \textbf{a}.An example of measured conductivity in LCR as a
    function of time. The slope of conductivity-time curve when
    $V_{\mathrm{G}}>0$ is interpolated, is smaller than that without
    gating. \textbf{b}. Molar diffusive flux through bare PCTE (blue
    bars: experimental; red bars: model values) and PG+PCTE (green
    bars) membranes for different salts. \textbf{c}. Experimentally
    measured salt rejection factor $\xi$ as a function of
    $V_{\mathrm{G}}$ for 0.1 mM KCl solution in HCR, showing an
    asymmetric response, with a higher degree of salt rejection for
    $V_{\mathrm{G}}>0$. \textbf{d}. $\xi_{\mathrm{max}}$ as a function
    of KCl concentration, fitted by a power law relation, suggesting
    the effect of Debye length on the observed salt rejection, since
    $\lambda_{\mathrm{D}} \propto c_{0}^{-\frac{1}{2}}$.}
  \label{fig:2} 
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/fig3.pdf}
  \caption{\textbf{Interplay between graphene quantum capacitance and
      EDL.} \textbf{a}. Schematic diagram of the electric
    potential at the graphene-electrolyte interface considering the
    quantum capacitance of graphene. The interfacial potential of the
    electrolyte solution, $\psi_{0}$, is smaller than the bias
    $V_{\mathrm{G}}$ applied, as a result of the change of graphene's
    Fermi level $\Delta \phi_{\mathrm{G}}$ and the potential drop in
    the Stern layer $\Delta \psi$. \textbf{b}. Calculated surface
    charge $\sigma_{\mathrm{G}}$ and surface potential
    $\psi_{\mathrm{G}}$ of graphene as a function of $V_{\mathrm{G}}$
    in a 1 mM KCl solution. }
  \label{fig:3}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/fig4.pdf}
  \caption{\textbf{Calculations using the proposed self-consistent
      theory}. \textbf{a}. Calculated cross-sectional contour plot of
    electric potential $\psi$ near the center of the graphene
    nanopore at $V_{\mathrm{G}}$ = 0.75 V. \textbf{b}. The
    corresponding electrochemical potential change for cation
    $\Delta \mu_{+}$ (left) and anion $\Delta \mu_{-}$ (right). The
    reference of electrochemical potential is set in the HCR far from
    the graphene layer. The diffusive pathways for both ions are
    indicated by the arrows. \textbf{c}. Calculated z-component of
    diffusive fluxes $\boldsymbol{J}_{z}$ for cation (left) and anion
    (right) through the pore as functions of the relative position
    inside the pore, considering different levels of $V_{\mathrm{G}}$
    applied. \textbf{d}. Calculated z-components of cation, anion, and
    total fluxes through the graphene nanopore as functions of
    $V_{\mathrm{G}}$, showing salt rejection for all fluxes.}
  \label{fig:4}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/fig5.pdf}
  \caption{\textbf{The Debye length effect on the degree of salt
      rejection}. \textbf{a}. Calculated contour plot of the surface
    potential of graphene $\psi_{\mathrm{G}}$ as a function of
    $\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ and $V_{\mathrm{G}}$. When
    the Debye length reduces, a higher $V_{\mathrm{G}}$ is required to
    maintain the same surface $\psi_{\mathrm{G}}$
    level. \textbf{b}. Calculated 2D contour plot of theoretically
    predicted $\xi$ as a function of
    $\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ and
    $V_{\mathrm{G}}$. Clearly, a higher degree of salt rejection can
    be achieved by having a higher $V_{\mathrm{G}}$ and
    $\lambda_{\mathrm{D}}/r_{\mathrm{G}}$}
  \label{fig:5}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/fig6.pdf}
  \caption{\textbf{Comparison between experimental and theoretical
      salt rejection considering different salt
      species}. \textbf{a}. Experimentally measured $\xi$ as a
    function of $V_{\mathrm{G}}$ for various salts at 0.1 mM. The
    error bar represents the standard error. \textbf{b}. Theoretically
    calculated $\xi-V_{\mathrm{G}}$ relations for the salt species
    considered here. \textbf{c}. Comparison between the experimental
    and calculated $\xi_{\mathrm{max}}$ values for different salt
    species, decreasing with the salt
    valence. \textbf{d}. $\xi_{\mathrm{max}}$ as a function of
    $\lambda_{\mathrm{D}}$ for the experimental (circle) and model
    (diamonds) values, showing a linear correlation.}
  \label{fig:6}
\end{figure}

\end{document}